---
title: "Homework 5"
author: "Ji-Eun(Jenn) Park"
date: "11/4/2017"
output:
  pdf_document:
    fig_height: 4.5
    fig_width: 4.5
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=TRUE, warning=TRUE,eval=TRUE,echo=TRUE,fig.pos='H',tidy=TRUE, tidy.opts=list(width.cutoff=60))
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_bw())
```

```{r cars}
summary(cars)
```

Question3
```{r setupdata, echo=FALSE}
setwd('~/Desktop/Biostatistics/HW')
dat <- read.csv('plco_dat_n25.csv')
```
• Part A: Is PSA at year 0 diﬀerent in men over 65 vs less than or equal to 65?
```{r}
dat %>%
  filter(age>65) %>%
  select(psa_level0) -> psa0_over

dat%>%
  filter(age<=65) %>%
  select(psa_level0) -> psa0_le
  
psa0_over <- as.numeric(psa0_over)

mean(psa0_over,na.rm=TRUE)
mean(table(psa0_le))
```

• Part B: Estimate the proportion of men with an abnormal DRE at year 0 (suspicious or not) among old and young men (including conﬁdence intervals). What can we conclude? Consult the data dictionary for the code associated with a suspicious DRE, and be sure to remove patients that are missing DRE screening that year.

• Part C: On average, is there a change in PSA from year 1 to year 2?

• Bonus: On average, is the change from year 0 to year 1 diﬀerent from the change from year 1 to year 2?

Add a plot to graphically show the data, and be sure to state the number of patients being used for this test.
