---
title: "Homework6"
author: "Ji-Eun(Jenn) Park"
date: "11/11/2017"
output:
  word_document: default
  pdf_document:
    fig_height: 4.5
    fig_width: 4.5
---

```{r options, include = FALSE}
knitr::opts_chunk$set(cache=TRUE, message=TRUE, warning=FALSE,eval=TRUE,echo=TRUE,fig.pos='H',tidy=TRUE, tidy.opts=list(width.cutoff=60))
library(tidyverse)
library(gridExtra) # present plots side-by-side
theme_set(theme_bw())
```

```{r q1setup, echo = FALSE, eval = FALSE}
install.packages("datasets")
library(datasets)
```

# Question 1 

Load the package datasets, which contains data on ﬂowers called iris. Use a permutation test to test whether or not Sepal Width and length are linearly associated, as measured by the correlation.

```{r q1, echo = FALSE}
(org_corr <- cor(iris$Sepal.Length,iris$Sepal.Width))

n <- nrow(iris)
B <- 10000

set.seed(1)
perm <- replicate(B, {
  perm_l <- sample(iris$Sepal.Length, size=n, replace=FALSE)
  perm_c <- cor(perm_l, iris$Sepal.Width)
})

(pvalue <- sum( abs(perm)  >  abs(org_corr)) / B )

```
Here we have used the permutation test to find out whether the length and width of Sepal is linearly associated. 

In this text, the null hypothesis is that the are not linearly associated,and the alternative hypothesis is that the two are linearly associated. Let's put the alpha level as 0.05. We consider that if the correlation of the permutation test is greater than the observed correlation, it is extreme and by using that we can find out that the p-value is 0.156. This is much bigger than our alpha level 0.05, so we fail to reject the null hypothesis. 

In other words, there is no significant evidence that there is a linear relationship between the width and length of Sepal.


# Question 2

In the United States, the same procedure may cost you much more in one hospital than in another one. In an October 27 article in the Portland Press Herald, the journalist Craig Anderson explains the issue and a new dataset available to compare prices across the state of Maine. The table below is from this article. 

Use a non-parametric test to compare the location of the distribution of prices of Maine Medical Center versus Central Maine Medical Center. Ignore the other hospitals.


```{r q2, echo = FALSE}
hosp <- matrix(c(34111,1258,87,656,14670,24789,210,
      30025,2618,22,1176,13655,29294,175), ncol=2)

x <- as.vector(hosp[,1])
y <- as.vector(hosp[,2])
df <- data.frame(X = x, Y = y)

df %>%
  mutate(Diff = X - Y) %>%
  mutate(Rank=rank(abs(Diff))) %>%
  filter(Diff > 0) %>%
  summarize(W = sum(Rank))

wilcox.test(x, y, alternative = "two.sided",paired = TRUE)
```
Here we want to test the difference of two samples, each observation related to each one in the other, so we use the Wilcoxon signed rank test. 

The null hypothesis would be the distribution of both population are equal. The alternative hypothesis would be that the distribution of the two are not equal, in other words, there is a  location shift between the two populations(either right or left). This shows that this will be a two-sided test. Let's put the alpha level as 0.05. The test statistic here is W statistic which is 13 in our test. Under the 0.05 significance level, our p-value here is 0.9375 which means we cannot reject the null hypothesis. 

In other words,a probability which would normally occur about everytime under the null hypothesis (Mann-Whitney U=53, p=1), not rare enough to suggest a statistically significant difference in ranked distributions between groups (given sig>0.05, 2-tailed). It can thus cannot be inferred that both Medical Centers are different.




# Question 3

A patient lives about 30 blocks from the hospital. She can either walk or take the bus to her appointments there, and it takes her on average the same time (30 minutes). However, the bus sometimes is faster and sometimes slower, so the standard deviation in her commute when she takes the bus is 7. She is more consistent when she walks, so the standard deviation of the time it takes her when she walks is only 2. Assume both times are normally distributed, with means and standard deviations as stated.

• Part A: Generate a random sample of 30 times when she walks, and another 30 times when she takes public transportation.
```{r q3a, echo = FALSE}
set.seed(2)
(walk <- rnorm(30,mean=30,sd=2))
(bus <- rnorm(30,mean=30,sd=7))
```

• Part B: Plot the empirical cumulative distribution function of each of these methods in the same plot.
```{r q3b, echo = FALSE}
e1 <- ecdf(walk)
e2 <- ecdf(bus)

plot(e1, xlim=c(min(walk, bus), max(walk, bus)), main="Illustration of KS test")
plot(e2, col=2, add=TRUE)
xy <- sort(c(walk, bus))
x.coord <- xy[which.max(abs(e1(xy) - e2(xy)))]
y.coord <- c(e1(x.coord), e2(x.coord))
segments(x.coord, y.coord[1], x.coord, y.coord[2], col=4, lwd=2, lty=2)
legend("topleft", c("sup |F(x) - G(x)|"), lwd=2, lty=2, col=4)
```

• Part C: Use a hypothesis test to compare the cumulative distribution function of these two methods of transportation.
```{r q3c, echo = FALSE}
ks.test(walk, bus, alternative = "two.sided")
```
Here we want to see the difference of spread of two ways to get to the hospital. Thus, we use a Kolmogorov-Smirnov test. 

The null hypothesis here is that time taken when walking and taking a bus has a same distribution. The alternative hypothesis is that they are not from the same distribution. Let's put the alpha level as 0.05. The test statistic D is 0.4 and when tested the hypotheses, the p-value is 0.016 which is p<0.05. Therefore, we can reject the null hypothesis. 

In other words, we have significant evidence that the time taking when walking and taking a bus is differently distributed.
# 뭐가 더 유용한지까지 결과까지 쓰기.


# Question 4

Imagine we wanted to study the eﬀect of caﬀeine on the time it took students to complete an exam. We design a randomized experiment where 50% of the students take a caﬀeine pill, while the rest take a placebo pill. 30 students are randomized and complete the exam in the following times:

% latex table generated in R 3.4.1 by xtable 1.8-2 package % Fri Nov 10 10:42:50 2017
```{r q4setup, echo = FALSE}
# Made a csv file with the same data with the table. 
# Import.
q4 <- read.csv('~/Desktop/datasets/question4.csv')
q4
```



• Part A: Assuming we know the treatment will not aﬀect the variance of the times (i.e. assume equal though unknown variances), compare the mean of the treatment arms at a 5% level, assuming times are normally distributed. Also compute a 95% conﬁdence interval for the diﬀerence in times.
```{r q4a, echo = FALSE}
# Students who took caffeine.
caffeine <- q4 %>%
            filter(treatment == 'caﬀeine')

# Students who did not take caffeine.
placebo <- q4 %>%
            filter(treatment == 'placebo')

# Assume normal distribution. Use t-test to compare.
t.test(caffeine$time_min, placebo$time_min, var.equal = TRUE)
```
Here we are comparing two independent samples and we are assuming they are both normally distributed with unknown but equal variance. Therefore, we can use a t-test with equal variance. 

The null hypothesis is that there is no difference between time taken to complete the test of the control and treatment group. The alternative hypothesis is that there is a difference. Let's put the alpha level as 0.05. The test statistics here is the t value which is -2.046 and the degree of freedom is 28. The 95% confidence interval is approximately from -15.88 to 0.01. The p-value is 0.05026 which is slightly above our significance level which is 0.05, so we cannot reject the null hypothesis. 

In other words, the mean of the time taken to complete the test of the control group who had placebo and the treatment group who had caffeinee does not have a statistically significant difference.



• Part B: It turns out that, although you were not originally told so, each student took two similar tests, once after taking the caﬀeine pill and once on the placebo pill. The data are on Canvas (dataHomework6_caﬀeine.csv). The treatment for the ﬁrst experiment is labelled treatment_exp1 and the time it took the student under this treatment is time_min_exp1. Similarly, the treatment received and the time for the second experiment are labelled as treatment_exp2 and time_min_exp2, respectively. Compare again the treatment arms, with a hypothesis test. Also compute a 95% conﬁdence interval for the diﬀerence in time between one treatment vs the other. How does this conﬁdence interval compare to the one in part A?
```{r q4b, echo = FALSE}
q4b <- read.csv('~/Desktop/datasets/dataHomework6_caffeine.csv')

caf1 <- q4b %>%
  filter(treatment_exp1 == 'caffeine')


caf2 <- q4b %>%
  filter(treatment_exp2 == 'caffeine')

caf2 <- caf2[,c(1,4,5,2,3)]
colnames(caf2) <- c("studentID", "treatment_exp1", "time_min_exp1", "treatment_exp2", "time_min_exp2")

newq4b <- rbind(caf1, caf2, by= 'studentID')
newq4b <- newq4b[-31,]
newq4b$time_min_exp1 <- as.numeric(as.character(newq4b$time_min_exp1))
newq4b$time_min_exp2 <- as.numeric(as.character(newq4b$time_min_exp2))

t.test(newq4b$time_min_exp1,newq4b$time_min_exp2,paired=TRUE)
```
Here we are comparing a paired sample. Therefore, we can use a paired t-test. 

The null hypothesis is that there is no difference between time taken to complete the test whether or not they had caffeine intake. The alternative hypothesis is that there is a difference. Let's put the alpha level as 0.05. The test statistics here is the t value which is -2.4908 and the degree of freedom is 29. The 95% confidence interval is approximately from -9.83 to -0.97. The p-value is 0.01872 which is lower than our significance level which is 0.05, so we can reject the null hypothesis. 

In other words, the mean of the time taken to complete the test of the control group who had placebo and the treatment group who had caffeinee does have a statistically significant difference.

The confidence level becomes narrower than the one in part A and it does not contain 0 which goes along with the null hypothesis and our rejection. In part A, it contained 0 which also goes along with the null hypothesis and our failure to reject it. The confidence interval also becomes narrower because we have more information in part b and when we use a paired sample, there is less bias so we have better estimations.


## Appendix
### Question 1-Setup code
```{r appendix00, ref.label='q1setup', eval= FALSE}
```

### Question 1 code
```{r appendix01, ref.label='q1', eval= FALSE}
```

### Question 2 code
```{r appendix1, ref.label='q2', eval = FALSE}
```

### Question 3-a code
```{r appendix2, ref.label='q3a', eval = FALSE}
```

### Question 3-b code
```{r appendix3, ref.label='q3b', eval = FALSE}
```

### Question 3-c code
```{r appendix4, ref.label='q3c', eval = FALSE}
```

### Question 4 Setup code
```{r appendix5, ref.label='q4setup', eval = FALSE}
```

### Question 4-a code
```{r appendix6, ref.label='q4a', eval = FALSE}
```

### Question 4-b code
```{r appendix7, ref.label='q4b', eval = FALSE}
```













